---
title: "STAT 385 - Homework 5"
author: "Jia Lin Mei - jmei43"
date: "Due 11:59 PM, 3/24/2020"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***

## Dataset: 2015 Flight Delays and Cancellations Data

In this homework, we take a quick look at the 2015 Flight Delays and Cancellations Data provided by the U.S. Department of Transportation. This is a huge dataset availalbe on [Kaggle](https://www.kaggle.com/usdot/flight-delays#airports.csv). But for us, we will only take a look at flights **flying out** from O'Hare International Airport (ORD) in January, 2015.

### Load data

- I have filtered out the data specific to O'Hare and stored it in `ohare_jan.csv`. This filtered data is available at the URL: https://nkha149.github.io/stat385-sp2020/files/data/ohare_jan.csv.

```{r message=FALSE}
library(tidyverse)
flights <- read_csv(file = "https://nkha149.github.io/stat385-sp2020/files/data/ohare_jan.csv")
```

- Write the code to print out the number of variables (columns) and the number of observations (rows) in this dataset.

```{r}
dim(flights)
#There are 26 variables and 23484 observations in this dataset
```

- Use the `View()` function to take a look at the data. (Don't add any code here)

***

### Review Basic Functions

First, let's review some basic R functions that we learned in the first half of the course.

Use R code to answer the following questions:

- How many different airlines fly out from O'Hare?

```{r}
newairlines <- unique(flights$AIRLINE)
length(newairlines)
```

- How many different airports is O'Hare connected to? (flights coming out of O'Hare go to)

```{r}
newdest <- unique(flights$DESTINATION_AIRPORT)
length(newdest)
```

- What is the average **departing delay** of flights departing O'Hare in Jan 2015?

```{r}
mean(flights$DEPARTURE_DELAY, na.rm=TRUE)
```

- What is the five summary statistics of the **taxi out time** of flights departing O'Hare in Jan 2015?

```{r}
newtaxiout <- na.omit(flights$TAXI_OUT)
summary(newtaxiout)
```

***

### `filter` function

Now, we will practice the skills we recently learned in the `dplyr` package.

- Print out only the flights that are going to U of I Willard Airport, `CMI`.
  - Make sure to remove `eval = FALSE` after your write the code!

```{r}
flights %>% 
  filter(DESTINATION_AIRPORT == "CMI")
```

- Print out only the flights that are going to Willard Airport on the weekends.
  - Make sure to remove `eval = FALSE` after your write the code!

```{r}
flights %>% 
  filter(DAY_OF_WEEK == "6" | DAY_OF_WEEK== "7") %>%
  filter(DESTINATION_AIRPORT == "CMI")
```

- Print out only the flights that are going to Willard Airport on the weekends that **are scheduled to arrive before 8:00 PM**.
  - Make sure to remove `eval = FALSE` after your write the code!

```{r}
flights %>% 
  filter(DAY_OF_WEEK == "6" | DAY_OF_WEEK== "7") %>%
  filter(DESTINATION_AIRPORT == "CMI") %>% 
  filter(SCHEDULED_ARRIVAL <= 2000)
```

***

### `select` function

- Of all the flights, print out only the following columns: `DESTINATION`, `DAY_OF_WEEK`, `SCHEDULED_DEPARTURE`, `DEPARTURE_TIME`, `DEPARTURE_DELAY`, `SCHEDULED_ARRIVAL`, `ARRIVAL_TIME`, `ARRIVAL_DELAY`.
  - Make sure to remove `eval = FALSE` after your write the code!

```{r}
flights %>% 
  select(DESTINATION_AIRPORT,DAY_OF_WEEK,SCHEDULED_DEPARTURE,DEPARTURE_TIME,
         DEPARTURE_DELAY,SCHEDULED_ARRIVAL,ARRIVAL_TIME,ARRIVAL_DELAY)
```

- Of all the flights going to Willard Airport on the weekend, print out all the columns except the following ones: `AIRLINE_DELAY`, `SECURITY_DELAY`, `AIR_SYSTEM_DELAY`.
  - Make sure to remove `eval = FALSE` after your write the code!

```{r}
flights %>% 
  filter(DAY_OF_WEEK == "6" | DAY_OF_WEEK== "7") %>%
  filter(DESTINATION_AIRPORT == "CMI") %>%
  select(-AIRLINE_DELAY,-SECURITY_DELAY,-AIR_SYSTEM_DELAY)
```

***

### `mutate` function

- Add a column that is the the ratio of the total taxing time (`TAXI_IN` and `TAXI_OUT`) and the flying time (`ELAPSED_TIME`). Name this new coumn `TAXI_RATIO`.
  - Make sure to remove `eval = FALSE` after your write the code!

```{r}
flights %>% 
  mutate (TAXI_RATIO = (TAXI_IN+TAXI_OUT)/ELAPSED_TIME)
```

***

### `groupby` and `summarize` functions

- Find the average departure delay time by destination and day of the week.
  - Make sure to remove `eval = FALSE` after your write the code!

```{r}
flights %>% 
  group_by(DESTINATION_AIRPORT,DAY_OF_WEEK) %>%
  summarize(ave_dep_delay_time = mean(DEPARTURE_DELAY, na.rm = TRUE), 
            n = n())
```

- Find the median taxi out time by airline and day of the week.
  - Make sure to remove `eval = FALSE` after your write the code!

```{r}
flights %>%
  group_by(AIRLINE,DAY_OF_WEEK) %>%
  summarize(med_taxiout_time = median(TAXI_OUT, na.rm = TRUE),
            n=n())
```

- Find the number of canceled flights for each airline.
  - Make sure to remove `eval = FALSE` after your write the code!

```{r}
flights %>% 
  group_by(AIRLINE) %>%
  summarise(canceled_flights = sum(CANCELLED == "1"))
```

- Find the ratio of canceled flights and the number of scheduled flights for each airline.
  - Make sure to remove `eval = FALSE` after your write the code!

```{r}
cancel <- flights %>% 
  group_by(AIRLINE) %>%
  summarise(canceled_flights = sum(CANCELLED == "1"),
            total_flights = n())
cancelratio <- cancel %>%
  mutate(cancel_ratio = canceled_flights/total_flights)
cancelratio
```


***

### `arrange` function

- Of the airlines that have at least 1000 scheduled flights, find the airline with the best canceling ratio record.
  - Make sure to remove `eval = FALSE` after your write the code!
```{r}
cancelratio %>% 
  arrange(desc(cancel_ratio))
#Airline MQ has the highest (best?) cancelling ratio.
```




